\documentclass{article}
\usepackage[sexy, hdr, fancy]{evan}
\setlength{\droptitle}{-4em}

\usepackage{tikz}
\usetikzlibrary{automata, arrows}

\lhead{Homework 3}
\rhead{Automata and Computation Theory}
\lfoot{}
\cfoot{\thepage}

\begin{document}
\title{Homework 3}
\maketitle
\thispagestyle{fancy}

\begin{enumerate}
	\item Give an NFA (both a state diagram and a formal description) recognizing the language $0^* 1^* 0^+$ with three states. The alphabet is $\left\{ 0, 1 \right\}.$
		\begin{soln}	
			The state diagram is given below:
			\begin{center}
				\begin{tikzpicture}[->, >=stealth', shorten >=1pt, auto, node distance=2cm, thick, main node/.style={circle, draw, font=\sffamily\Large\bfseries}]

					\node[initial, state] (0) {$q_0$};
					\node[state] (1) [right of = 0] {$q_1$};
					\node[state, accepting] (2) [right of = 1]{$q_2$};

					\path[every node/.style={font=\sffamily\small}]
					(0) edge [loop above] node {0} (0)
					edge node {$\varepsilon$} (1)
					(1) edge [loop above] node {1} (1)
					edge node {0} (2)
					(2) edge [loop above] node {0} (2);
				\end{tikzpicture}
			\end{center}

			The states are $Q=\left\{ q_0, q_1, q_2 \right\},$ the start state is $q_0,$ the accept states are $F=\left\{ q_2 \right\},$ the alphabet is $\Sigma=\left\{ 0, 1 \right\},$ and the transition function is given by
			\begin{center}
				\begin{tabular}{c|ccc}
					$\delta$ & 0 & 1 & $\varepsilon$ \\
					\hline
					$q_0$ & $\left\{ q_0 \right\}$ & $\varnothing$ & $\left\{ q_1 \right\}$ \\
					$q_1$ & $\left\{ q_2 \right\}$ & $\left\{ q_1 \right\}$ & $\varnothing$ \\
					$q_2$ & $\left\{ q_2 \right\}$ & $\varnothing$ & $\varnothing$
				\end{tabular}
			\end{center}
		\end{soln}

		\newpage
	\item This question studies the number of states in a DFA equivalent to an NFA. Recall that in class we showed an NFA with 4 states that recognizes the language which consists of all binary strings that have a 1 in the third position from the end. For any integer $k,$ it is easy to generalize this construction to an NFA with $k+1$ states that recognizes the language which consists of all binary strings that have a 1 in the $k$th position from the end. The general transformation from an NFA too a DFA will give us a DFA with at most $2^{k+1}$ states recognizing the same language.

		Show that, any DFA that recognizes the same language must have at least $2^k$ states.

		Hint: start by looking at the following two strings: $10^{k-1}$ and $0^k.$ Observe that when a DFA takes them as inputs, it must end up at different states, since one string is accepted and the other is rejected.
		\begin{proof}
			Fix $k$ and consider two strings $a=a_1a_2\cdots a_k$ and $b=b_1b_2\cdots b_k$ of length $k$ that are different. Then there must exist an index $i$ where they differ. WLOG $a_i=1$ and $b_i=0.$ Now let $c=0^{i-1}.$ Then $ac$ is a string where the $k$th position from the end is a 1, and $bc$ is a string where the $k$th position from the end is a 0. Thus, $ac$ would be accepted while $bc$ is not, so these strings must be different states.

			Thus, for any two strings of length $k,$ there must be two distinct states in the DFA to account for the above process. Since there are $2^k$ strings of length $k,$ there must be at least $2^k$ distinct states in the DFA.
		\end{proof}

		\newpage
	\item Say that string $x$ is a prefix of string $y$ if a string $z$ exists where $xz=y$ and that $x$ is a proper prefix of $y$ if in addition $x\neq y.$ Let $A$ be a regular language. Show that the class of regular languages is closed under the following operation.
		\begin{align*}
			\text{NOEXTEND}(A) = \left\{ w\in A:w\text{ is not the proper prefix of any string in }A \right\}
		\end{align*}

		Hint: Think about when a string $w\in A$ can be the proper prefix of another string in $A,$ then modify the states of the machine to avoid this.
		\begin{proof}
			Let $M=(Q, \Sigma, \delta, q_0, F)$ be the DFA recognizing $A,$ since $A$ is regular. We wish to construct an DFA $N$ that accepts NOEXTEND$(A).$ 

			Consider a string $w$ that reaches an accept state $q\in F.$ Then $w\in A.$ If there exists a string $x$ that reaches $p\in F$ from $q,$ we know that $w$ is the proper prefix of a string in $A$ since $wx$ is accepted. This path can be detected using a DFA. Now, let $F'\subset F$ be the subset of $F$ such that there are no strings from $F'$ to $F.$ Then $N=(Q, \Sigma, \delta, q_0, F')$ is a DFA.

			If $w\in\text{NOEXTEND}(A),$ then $w\in A$ so $w$ is a path to some state in $F,$ and this state is in $F'$ because $w$ is not the proper prefix of any other accepted string.

			Conversely, if $w$ is accepted by $N,$ it by construction it must have been accepted by $M$, so $w\in A,$ and $w$ is only accepted if there is no path to another state in $F,$ so if $w$ is not the proper prefix of a string in $A.$ Thus, $N$ recognizes NOEXTEND$(A),$ so it is regular.
		\end{proof}

		\newpage
	\item Let $\Sigma=\left\{ 0, 1 \right\}.$
		\begin{enumerate}[(a)]
			\item Write a regular expression for the language $L$ consisting of all strings in $\Sigma^*$ with exactly one occurrence of the substring 000.
				\begin{soln}
					The string 000 must be immediately surrounded by arbitrary$(1^+0)$ and $(1^+00)$ on the right, and $(01^+)$ and $(001^+)$ on the left, if there is anything to the left or right, respectively. Then since the string could start or end with arbitrary 1s, the regular expression is
					\begin{align*}
						1^*\left[ (001^+)\cup(01^+) \right]^*000\left[ (1^+0)\cup(1^+00) \right]^*1^*
					\end{align*}
				\end{soln}

			\item Write a regular expression for the language $L$ consisting of all strings in $\Sigma^*$ that do not end with 00.
				\begin{soln}
					The string can have length 0 or 1. Otherwise, the string can only end in 01, 10, or 11, and the rest can be anything else, so the regular expression is
					\begin{align*}
						(0\cup 1)^* (01\cup 11\cup 10)\cup(0\cup1\cup\varepsilon)
					\end{align*}
				\end{soln}

		\end{enumerate}

\end{enumerate}

\end{document}
