\documentclass{article}
\usepackage[sexy, hdr, fancy]{evan}
\setlength{\droptitle}{-4em}

\lhead{Homework 5}
\rhead{Introduction to Financial Derivatives}
\lfoot{}
\cfoot{\thepage}

\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}

\begin{document}
\title{Homework 5}
\maketitle
\thispagestyle{fancy}

\section*{Chapter 6: Interest Rate Futures}

\begin{itemize}
	\item[4.] A Eurodollar futures price changes from 96.76 to 96.82. What is the gain or loss to an investor who is long two contracts?

	\item[6.] The 350-day LIBOR rate is 3\% with continuous compounding and the forward rate calculated from a Eurodollar futures contract that matures in 350 days is 3.2\% with continuous compounding. Estimate the 440-day zero rate.

	\item[9.] It is May 5, 2014. The quoted price of a government bond with a 12\% coupon that matures on July 27, 2024, is 110-17. What is the cash price?

	\item[11.] It is July 30, 2015. The cheapest-to-deliver bond in a September 2015 Treasury bond futures contract is a 13\% coupon bond, and delivery is expected to be on September 30, 2015. Coupon payments on the bond are made on February 4 and August 4 each year. The term structure is flat, and the rate of interest with semiannual compounding is 12\% per annum. The conversion factor for the bond is 1.5. The current quoted bond price is \$110. Calculate the quoted futures price for the contract.

	\item[14.] Suppose that the 300-day LIBOR zero rate is 4\% and the Eurodollar quotes for contracts maturing in 300, 398, and 489 days are 95.83, 95.62, and 95.48. Calculate 398-day and 489-day LIBOR zero rates. Assume no difference between forward and futures rates for the purposes of your calculation.

	\item[21.] The 3-month Eurodollar futures price for a contract maturing in 6 year is quoted as 95.20. The standard deviation of the change in the short term interest rate in 1 year is 1.1\%. Estimate the forward LIBOR interest rate for the period between 6.00 and 6.25 years in the future.

	\item[26.] A Eurodollar futures quote for the period between 5.1 and 5.35 years in the future is 97.1. The standard deviation of the change in the short-term interest rate in one year is 1.4\%. Estimate the forward interest rate in an FRA.
		
	\item[27.] It is March 10, 2014. The cheapest-to-deliver bond in a December 2014 Treasury bond futures contract is an 8\% coupon bond, and the delivery is expected to be made on December 31, 2014. Coupon payments on the bond are made on March 1 and September 1 each year. The rate of interest with continuous compounding is 5\% per annum for all maturities. The conversion factor for the bond is 1.2191. The current quoted bond price is \$137. Calculate the quoted futures price for the contract.
		
\end{itemize}

\end{document}
